---
title: "Simulation"
author: "Amy Lee"
date: "November 8, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Simulation Project
In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem.   The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter.  The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.  Set lambda = 0.2 for all of the simulations.  You will investigate the distribution of averages of 40 exponentials.  Note that you will need to do a thousand simulations.  


Illustrate via simulation and associated explanatory text the properties of the distribution of the mean of 40 exponentials. You should

1) Show the sample mean and compare it to the theoretical mean of the distribution.
2) Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.
3) Show that the distribution is approximately normal.

---

# Review Criteria
- Did you show where the distribution is centered at and compare it to the theoretical center of the distribution?
- Did you show how variable it is and compare it to the theoretical variance of the distribution?
- Did you perform an exploratory data analysis of at least a single plot or table highlighting basic features of the data?
- Did the student perform some relevant confidence intervals and/or tests?
- Were the results of the tests and/or intervals interpreted in the context of the problem correctly?
- Did the student describe the assumptions needed for their conclusions?

---

# Overview

This project was completed to observe the effect of Central Limit Theorem.
Central Limit Theorem states, the distribution of the sample means approximate a normal distribution as the sample size increases. 
In this project, we will use the exponential distribution as our sample where lambda=0.2 and each sample size=40.

Before we start the project, packages must be imported first.

#### Necessary packages
```{r message=FALSE}
library(ggplot2)
library(Hmisc)
```

And now, we need to make the samples (1000 samples of n=40, lambda=0.2 exp distributions)

#### Making Samples
```{r}
exp<-c()
for (i in 1:1000){
  exp<-rbind(exp,rexp(40,0.2))
}
head(exp)
```

Here is the sample data. Each row is one sample exponential distribution with lambda = 0.2 and n = 40.
There are 1000 sample datas (1000 rows)
Now, we can start answering question.

## 1) Show the sample mean and compare it to the theoretical mean of the distribution.


```{r}
mns<-c()
for (i in 1:1000) {
  mns[i]<-mean(exp[i,])
}
means<-cumsum(mns[1:1000])/(1:1000)
```

First, I made the mns vector which contains the means of all distribution of the 1000 samples.
And made the cumulative sum vector called means.

```{r}
vrs<-c()
for (i in 1:1000) {
  vrs[i]<-var(exp[i,])
}
p<-ggplot(data.frame(x=1:1000, y=mns), aes(x=x, y=y))+geom_point()+geom_line(size=0.5)
p<-p+ylim(min(mns-qnorm(0.975)*sqrt(vrs/40))-0.1,max(mns+qnorm(0.975)*sqrt(vrs/40))+0.1)+geom_hline(yintercept=5, size=1, color="red")
p<-p+labs(title="Comparing True Mean to the Each Sample Mean", x="Sample Number", y="Mean")
p<-p+geom_pointrange(aes(ymin=mns-qnorm(0.975)*sqrt(vrs/40), ymax=mns+qnorm(0.975)*sqrt(vrs/40)), color=ifelse(mns-qnorm(0.975)*sqrt(vrs/40)>5 | mns+qnorm(0.975)*sqrt(vrs/40)<5, "red","black"), alpha=0.2)
p
```

The graph above shows the mean of each 1000 samples and its 95% interval. 
If the confidence interval did not contain the true mean of 5, it was colored red. 
The graph above doesn't clearly show the CI, so I will show the first 50 samples below.

```{r}
p1<-ggplot(data.frame(x=1:50, y=mns[1:50]), aes(x=x, y=y))+geom_point()+geom_line(size=0.5)
p1<-p1+ylim(min(mns[1:50]-qnorm(0.975)*sqrt(vrs[1:50]/40))-0.1,max(mns[1:50]+qnorm(0.975)*sqrt(vrs[1:50]/40))+0.1)+geom_hline(yintercept=5, size=1, color="red")
p1<-p1+labs(title="Comparing True Mean to the Each Sample Mean (50 samples)", x="Sample Number", y="Mean")
p1<-p1+geom_pointrange(aes(ymin=mns[1:50]-qnorm(0.975)*sqrt(vrs[1:50]/40), ymax=mns[1:50]+qnorm(0.975)*sqrt(vrs[1:50]/40)), color=ifelse(mns[1:50]-qnorm(0.975)*sqrt(vrs[1:50]/40)>5 | mns[1:50]+qnorm(0.975)*sqrt(vrs[1:50]/40)<5, "red","black"), alpha=0.2)
p1
```






Then, I calculated the cumulative means to see how the mean changes as the sample size increased.

```{r}
g_m<-ggplot(data.frame(x=1:1000, y=means), aes(x=x, y=y))+geom_line(size=0.5)+geom_point()
g_m<-g_m+ylim(min(means)-0.1,max(means)+0.1)+geom_hline(yintercept=5, size=1, color="red")
g_m<-g_m+labs(title="Change in Sample Mean as Sample Size Increase", x="Sample Size", y="Mean")
g_m
```

As it can be observed on the graph, the sample mean converges to the theoretical mean of 5.

## 2) Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

Variance of the sample mean is the variance of the population over the sample size.
So, the variance of the exponential distribution is 1/lambda^2 which is 25.
And the sample size is 40. This makes the variance of the sample mean 0.625
So the variance of the sample should be similar to 0.625.

```{r}
var(mns)
```

As shown above, it can be seen that the sample variance is similar to the theoretical variance.

3) Show that the distribution is approximately normal.
